### 2024-04-22 paper/paper-part2.tex
refactor literature search for #3

\section{Related Work}

Why have we choosen model X and model Y for our work?

From~\cite{jiao20}: Language model pre-training, such as BERT, has significantly improved the performances of many natural language processing tasks. However, pre-trained language models are usually computationally expensive, so it is difficult to efficiently execute them on resource-restricted devices. To accelerate inference and reduce model size while maintaining accuracy, we first propose a novel Transformer distillation method that is specially designed for knowledge distillation (KD) of the Transformer-based models. By leveraging this new KD method, the plenty of knowledge encoded in a large teacher BERT can be effectively transferred to a small student Tiny-BERT. Then, we introduce a new two-stage learning framework for TinyBERT, which performs Transformer distillation at both the pretraining and task-specific learning stages. This framework ensures that TinyBERT can capture he general-domain as well as the task-specific knowledge in BERT.
TinyBERT with 4 layers is empirically effective and achieves more than 96.8\% the performance of its teacher BERTBASE on GLUE benchmark, while being 7.5x smaller and 9.4x faster on inference. TinyBERT with 4 layers is also significantly better than 4-layer state-of-the-art baselines on BERT distillation, with only about 28\% parameters and about 31\% inference time of them. Moreover, TinyBERT with 6 layers performs on-par with its teacher BERTBASE.

Examples further include ``production ready systems'' like NBoost for ElasticSearch.\footnote{\url{https://github.com/koursaros-ai/nboost}}

- NBoost is a scalable, search-api-boosting platform for deploying
transformer models to improve the relevance of search results or
diffe rent platforms (i.e. Elasticsearch) Topics
- maybe more useful for the user pipeline
- is between user and the actual pipeline where it picks the best
based on relevance and training

JMC and DISP~\cite{cui21}:
- Joint Model Compression (JMC)
- Direct Importance-aware Structured Pruning
- Dense Knowledge Distillation (DKD)
- Paper at thulb

MiniLMv2~\cite{wang21} (https://huggingface.co/sentence-transformers/all-MiniLM-L6-v2)
- found in https://arxiv.org/abs/2012.15828 (MiniLMv2)
- they state that their model outperform state of the art models
- uses deep distillated self-attention which are distributed and also
give values to relations to copy teacher self-attention
- transfers knowledge of upper-middle- to last-layer(MiniLM only
mimics the last layer)

Transit(https://arxiv.org/abs/2101.03289)