### 2024-05-23 paper/cross-encoder.tex
doc: how we choosed hyperparameters and results #7

To determine the appropriate hyperparameters for our training system, we systematically evaluated various parameters and the performance of the resulting cross-encoder models. This method ensures that the model achieves a high Mean Reciprocal Rank (recip\_rank) and a high Normalized Discounted Cumulative Gain (ndcg\_cut.10). Our evaluation focused on four primary parameters: batch size, learning rate, and the effect of randomly initialized weight matrices.

Initially, hyperparameters were intuitively selected and designated as the baseline model, reflecting the default values of the training method employed. To explore alternative hyperparameters, we tested values both below and above the defaults. For each parameter to be fine-tuned, three different values were selected. Each parameter was altered individually with respect to the baseline, resulting in the generation, training, and evaluation of 13 models in total. This approach ensures a comprehensive assessment of each hyperparameter's impact on model performance.

\begin{table}[h!]
\centering
\begin{tabular}{|c|c|c|c|c|c|c|c|}
\hline
index & name & bs & epochs & lr & seed & ndcg\_cut.10 & recip\_rank \\
\hline
0 & BM25 & & & & & 0.4795 & 0.7950 \\
\hline
\multicolumn{8}{|c|}{base line:} \\
\hline
1 & cross-encoder- & 32 & 10 & 2e-5 & 233 & 0.6251 & 0.9162 \\
\hline
\multicolumn{8}{|c|}{batch size:} \\
\hline
2 & cross-encoder- & 16 & 10 & 2e-5 & 233 & 0.6045 & 0.8901 \\
3 & cross-encoder- & 64 & 10 & 2e-5 & 233 & 0.6309 & 0.9147 \\
4 & cross-encoder- & 128 & 10 & 2e-5 & 233 & 0.6248 & 0.8904 \\
\hline
\multicolumn{8}{|c|}{epochs:} \\
\hline
5 & cross-encoder- & 32 & 5 & 2e-5 & 233 & 0.6100 & 0.9036 \\
6 & cross-encoder- & 32 & 15 & 2e-5 & 233 & 0.6380 & 0.9005 \\
7 & cross-encoder- & 32 & 20 & 2e-5 & 233 & 0.6319 & 0.9009 \\
\hline
\multicolumn{8}{|c|}{learning rate} \\
\hline
8 & cross-encoder- & 32 & 10 & 1e-5 & 233 & 0.6040 & 0.9152 \\
9 & cross-encoder- & 32 & 10 & 1e-4 & 233 & 0.5832 & 0.8843 \\
10 & cross-encoder- & 32 & 10 & 5e-4 & 233 & 0.5129 & 0.8502 \\
\hline
\multicolumn{8}{|c|}{seed:} \\
\hline
11 & cross-encoder- & 32 & 10 & 2e-5 & 5 & 0.6241 & 0.9044 \\
12 & cross-encoder- & 32 & 10 & 2e-5 & 6765 & 0.6170 & 0.9068 \\
13 & cross-encoder- & 32 & 10 & 2e-5 & 46368 & 0.6200 & 0.8947 \\
\hline
\end{tabular}
\caption{Evaluation of Cross-Encoder Models with Various Hyperparameters}
\label{table:cross_encoder}
\end{table}