### 2024-06-24 paper/paper-part3.tex
mf

\section{TinyBERT for Passage Re-Ranking on MS~MARCO}

\subsection{Plain TinyBERT}

\subsection{Distilling RankZephyr into TinyBERT}

\subsection{BM25 Score Injection for TinyBERT}